---
title: "CQC Data Package"
subtitle: "User Guide"
author: "Ritwika Das"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: false
    theme: cosmo
  gfm:
    toc: true
---

::: {.callout-note}

## Part 1: For Frontline Users

::: {.callout-important}

### For End Users

**You do NOT need to build, update, or maintain datasets.** This package uses GitHub Actions to automatically update dataframes daily from the CQC API. You can simply install the package from GitHub and access the automatically-updated provider, location or merged datasets. You can also use `get_cqc_id_info()` to query any CQC provider or location. The datasets cover both location-level data (individual care facilities and services) and provider-level data (organizations that run these services).

:::

### What is CQC?

The Care Quality Commission (CQC) is the independent regulator of health and social care services in England. It monitors, inspects, and rates services to ensure they meet fundamental standards of quality and safety. The CQC oversees a wide range of care providers including hospitals, GP practices, care homes, dental practices, ambulance services, and mental health services. Their ratings system (Outstanding, Good, Inadequate and Needs Improvement) helps the public make informed choices about their care while also attempting to encourage improvement across the sector.

### What Data is Available?

Once the package is loaded, three ready-to-use datasets are available immediately:

| Dataset | Description |
|---|---|
| `location_df` | Individual care facilities and services |
| `provider_df` | Organisations that run those services |
| `merged_df` | Combined dataset linking providers with their locations |

For most users, `merged_df` is the most useful as it contains both location and provider information in one place.

## Quick Start: What You Need to Know

### Installation

The cqcpack package is available on GitHub and comes with daily-updated datasets on providers, locations and a combination of the two:

```{r pack-install}
#| include: true
#| message: false
#| warning: false
#| eval: false

# Install remotes if you don't have it
install.packages("remotes")
  
# Install from GitHub
remotes::install_github("Care-Policy-and-Evaluation-Centre/cqcpack", force = TRUE)

# Load the package 
library(cqcpack)
```

**Note:** These datasets are pre-processed dataframes updated daily from the CQC API. When you load the package, you will always have access to data from within the last 24 hours.

::: {.callout-important}
**Windows Only.** This package has currently only been tested on Windows. Mac users may experience issues during installation where R and data files are cloned as blank files. We are investigating this and hope to add Mac support in a future release. If you are on a Mac and encounter this issue, please raise it on our [GitHub Issues page](https://github.com/Care-Policy-and-Evaluation-Centre/cqcpack/issues).
:::

### Accessing Pre-Built Datasets

The package comes with ready-to-use datasets. You can access them directly without any building or updating:

```{r pack-chck}
#| include: true
#| message: false
#| warning: false
#| eval: false

# Load your preferred dataset
merged_df
        
# Check the data 
dim(merged_df) 
head(merged_df, 5)  
```

### Analysing ratings in your area

```{r ratings-summary}
#| include: true
#| message: false
#| warning: false
#| eval: false

library(knitr)

# Summary of ratings
table(merged_df$location_currentRatings_overall_rating)

# Filter for Good-rated services
good_rated <- merged_df[merged_df$location_currentRatings_overall_rating == "Good", ]
head(good_rated, 5)

ratings <- as.data.frame(table(Rating = merged_df$location_currentRatings_overall_rating))
ratings$Percentage <- paste0(round(ratings$Freq / sum(ratings$Freq) * 100, 1), "%")
colnames(ratings) <- c("Rating", "Count", "Percentage")

knitr::kable(
  ratings,
  caption = "Breakdown of CQC overall ratings across all registered services",
  align = c("l", "r", "r"),
  format.args = list(big.mark = ",")
)
```

### Sample Data

The table below shows the first 6 rows of the merged dataset, displaying some relevant columns:

```{r data-visual}
#| include: true
#| message: false
#| warning: false
#| eval: false

# Select the most useful columns to display
sample_cols <- c(
  "location_name",
  "location_type",
  "location_region",
  "location_postalCode",
  "location_registrationStatus",
  "location_currentRatings_overall_rating",
  "provider_name"
)

kable(
  head(merged_df[, sample_cols], 6),
  col.names = c(
    "Location Name",
    "Type",
    "Region",
    "Postcode",
    "Status",
    "Overall Rating",
    "Provider Name"
  ),
  caption = "Sample of the CQC merged dataset"
)
```

### Querying CQC IDs

The primary function you will use is get_cqc_id_info(). This retrieves detailed information for any specific CQC provider or location ID.

```{r id-test}
#| include: true
#| message: false
#| warning: false
#| eval: false
        
# For a location ID
location_info <- get_cqc_id_info("1-10000302982", id_type = "location")

# For a provider ID
provider_info <- get_cqc_id_info("1-102642938", id_type = "provider")

print(location_info)
print(provider_info)
```

### Visualising Provider Size

As an example, I am interested in inspecting adult social care locations that fall under the social care organisation category type and also have beds.
```{r beds-data}
#| include: true
#| message: false
#| warning: false
#| eval: false

unique(merged_df$location_type) # eg. Primary Medical Services, Social Care Org, Independent Ambulance, etc.
table(merged_df$location_type[merged_df$location_numberOfBeds > 0])
summary(merged_df$location_numberOfBeds)

beds_data <- merged_df[
  merged_df$location_type == "Social Care Org" & 
  merged_df$location_inspectionDirectorate == "Adult social care" &
  merged_df$location_numberOfBeds > 0 &
  !is.na(merged_df$location_numberOfBeds),
]
summary(beds_data$location_numberOfBeds)
```

```{r ratings-visual}
#| include: true
#| message: false
#| warning: false
#| eval: false

library(ggplot2)

beds_data$size_category <- cut(
  beds_data$location_numberOfBeds,
  breaks = c(0, 10, 40, 100, 200, Inf),
  labels = c(
    "Small (1-10)", 
    "Medium (11-40)", 
    "Large (41-100)", 
    "Very Large (101-200)", 
    "Extra Large (200+)"
  ),
  right = TRUE
)

ggplot(beds_data, 
  aes(x = size_category, fill = location_currentRatings_overall_rating)) +
  geom_bar(position = "stack") +
  scale_y_continuous(labels = scales::comma) +
  scale_fill_manual(
    values = c(
      "Outstanding" = "#006747",
      "Good" = "#78BE20",
      "Requires improvement" = "#F5A623",
      "Inadequate" = "#DA1E28",
      "Inspected but not rated" = "#546E7A",
      "No published rating" = "#CFD8DC",
      "Insufficient evidence to rate" = "#9E9E9E"
    ),
    na.value = "#EEEEEE"
  ) +
  labs(
    title = "CQC Ratings by Location Size",
    subtitle = "Adult social care residential locations only",
    x = "Location Size",
    y = "Number of Locations",
    fill = "Overall Rating"
  ) +
  theme_minimal()
```

The graph demonstrates that a majority of social care locations operate with less than 100 beds. It also shows that "Good" is the most commonly received rating across location sizes. Relative to them, there are far fewer locations that receive "Inadequate" or event "Requires Improvement" ratings. There also seems to be quie a few locations with "NA" ratings which means they may be newly registered and/or their inspection may be in progress. 

```{r na-locations}
#| include: true
#| message: false
#| warning: false
#| eval: false

na_rated <- beds_data[is.na(beds_data$location_currentRatings_overall_rating), ]
na_dates <- as.Date(na_rated$location_registrationDate)
summary(na_dates)
```

However, upon closer inspecition, unrated providers span registration dates from 2010 to the present day, which suggests that missing ratings could be due to newly registered services awaiting their first inspection and also deregistered or closed services. It could also evidence historical data gaps in CQC records.

:::

::: {.callout-note}

## Part 2: Background Technical Information

::: {.callout-warning}

## Advanced Users Only

The following sections describe the internal workings of the package. **Regular users do not need to use these functions or maintain any data.** The package is automatically updated via GitHub Actions, and the data is pre-built and maintained for you.

:::

## How This Package Works

### Data Source

The package retrieves data directly from the official CQC API endpoints through an automated GitHub Actions workflow. The repository contains pre-processed dataframes that are automatically updated daily via scheduled runs. These dataframes include comprehensive datasets about providers and locations, with their registration details, ratings, inspection dates, and service characteristics.

The raw JSON data from the CQC API is stored in a separate repository: [cqc-data-repo](https://github.com/Care-Policy-and-Evaluation-Centre/cqc-data-repo.git), while the processed dataframes live in the package itself. This automation means users always have access to recent data without needing to call the API themselves or perform any maintenance.

### What this Package Covers

The `cqcpack` package provides a complete toolkit for accessing and managing CQC regulatory data. The package handles the full data lifecycle including initial bulk data collection, caching for efficiency, incremental updates to keep data current, and merging capabilities to create comprehensive datasets linking providers with their locations.

### Why this Approach?

Working directly with the CQC API requires handling pagination, rate limits, JSON parsing, and wrangling complex data structures. This package eliminates these challenges by providing functions that handle all the complexity behind the scenes. It offers:

-    Bulk data collection with local caching to avoid redundant API calls

-    Automated incremental updates that capture only changes since the last data pull

-    Built-in data transformation that converts nested JSON structures into analysis-ready dataframes

-    The ability to merge provider and location data for comprehensive analysis

## Technical Functions (for Package Maintainers)

### Caching Functions

These functions handle the initial data collection and storage:

-    `cache_location_ids()` - Downloads and caches all location identifiers

-    `cache_location_jsons()` - Fetches and stores detailed JSON data for all locations

-    `cache_provider_ids()` - Downloads and caches all provider identifiers

-    `cache_provider_jsons()` - Fetches and stores detailed JSON data for all providers

### Data Building Functions

These functions construct dataframes from cached data:

-    `build_location_df()` - Constructs a dataframe from cached location JSON files

-    `build_provider_df()` - Constructs a dataframe from cached provider JSON files

### Data Integration Functions

-    `merge_provider_location()` - Combines provider and location datasets into a unified dataframe

### Update Functions

These functions maintain currency of the datasets:

-    `get_incremental_changes()` - Identifies changes since the last data collection

-    `update_location_dataset()` - Updates location data with recent changes

-    `update_provider_dataset()` - Updates provider data with recent changes

### Query Functions

-    `get_cqc_id_info()` - Retrieves detailed information for specific provider or location IDs (PRIMARY USER FUNCTION)

## Automated Update System

The package uses GitHub Actions to automatically maintain current data:

1.   **Scheduled Updates**: Runs daily to check for new data from the CQC API

2.   **Incremental Updates**: Uses `get_incremental_changes()`, `update_location_dataset()`, and `update_provider_dataset()` to fetch only new or changed records

3.  **Data Storage**:

    -    Processed dataframes - package's `data/` folder

    -    Raw JSON files - [cqc-data-repo](https://github.com/Care-Policy-and-Evaluation-Centre/cqc-data-repo.git)

4.  **Automatic Deployment**: Updated dataframes are automatically committed to the package repository

To get the latest data, simply reinstall the package by running the install command again.

## Manual Update Workflow (Optional)

For troubleshooting or manual updates, here's the process that GitHub Actions runs automatically:

```         
library(cqcpack)  

# Get incremental changes from CQC API 
changes <- get_incremental_changes() 
str(changes, max.level = 2)  

# Update datasets with recent changes 
update_location_dataset() 
update_provider_dataset()  

# Create updated merged dataset 
merged_data <- merge_provider_location()  

# Verify the updates 
tail(merged_data[order(merged_data$date), c("locationName", "date")], 3) 
```
:::

## Summary

**For end users:** Install the package from GitHub and use `get_cqc_id_info()` or `merge_provider_location()`. The data is automatically kept up-to-date via GitHub Actions. That's it!

**For package maintainers:** GitHub Actions handles daily updates automatically. The workflow uses incremental update functions to fetch new data and stores processed dataframes in the package and raw JSONs in the [cqc-data-repo](https://github.com/Care-Policy-and-Evaluation-Centre/cqc-data-repo.git).
